---
title: "Create ConScape Inputs using 2021 HF data"
output:
  html_document:
    df_print: paged
---


```{r}
library(terra)
library(sf)
library(here)
library(dplyr)
library(tmap)
library(gdalUtilities)
library(fasterize)
```


## 1. Extract HF 2021 data for study area from gdb

### a) Connect terra to gdb using a proxy

Using a proxy means that the data are not read into memory.

```{r}
f <- "/Users/celia/Documents/ArcGIS/Projects/celia_HFI2021/HFI2021.gdb"
v <- vect(f, proxy=TRUE)
```
### b) Define extent of study area

- Import polygon of extent of area of interest (AOI) with the same crs as gdb

```{r}
AOI <- st_read(here::here("data/AOE_shp/Proposed Focus Area - connectivity.shp"))
```

Map the study area to double check

```{r}
tmap_mode("view")
tm_shape(AOI) + tm_borders()
```

### c) Extract HF data for study area

Extract data and convert to sf object 

```{r}
HF2021_SE <- query(v, extent=ext(AOI))
HF2021_SE <- st_as_sf(HF2021_SE)
HF2021_SE
```

### d) Clean HF data

Code from Viviana's project

```{r}
ensure_multipolygons <- function(X) {
    tmp1 <- tempfile(fileext = ".gpkg")
    tmp2 <- tempfile(fileext = ".gpkg")
    st_write(X, tmp1)
    ogr2ogr(tmp1, tmp2, f = "GPKG", nlt = "MULTIPOLYGON")
    Y <- st_read(tmp2)
    st_sf(st_drop_geometry(X), geom = st_geometry(Y))
}
```

```{r}
HF2021_SE <- ensure_multipolygons(HF2021_SE)
HF2021_SE <- st_make_valid(HF2021_SE)
```

### e) Export to .gpkg

```{r}
#st_write(HF2021_SE, here::here("data/HF2021_SE_CH.gpkg"), append=FALSE)
```

## 2. Add FU values

Read in HF2021_Kainai from .gpkg (sf object)

```{r}
HF2021_SE <- st_read(here::here("data/HF2021_SE.gpkg"))
```

Import FU values (Marrec et al. 2020)

```{r}
FU_values <- readxl::read_excel(here::here("data/FU_values.xlsx"))
#View(FU_values)
```

Check that HF categories match (I modified the list so that they match for this test file, and added category "NATURAL"). 

I couldn't find exact information on "HARVEST-AREA-WHITE-ZONE" but took an educated guess that it means a Forest Harvest area < 4 years old.
Also had a hard time finding appropriate info for "CAMP-INDUSTRIAL" and took a conservative approach and set it equal to "urban industrial"


```{r}
table(is.element(unique(HF2021_SE$FEATURE_TY), FU_values$FEATURE_TY))
```

Add F and U values to polygons based on their feature type. 

```{r}
HF2021_SE <- left_join(HF2021_SE, FU_values %>% select(FEATURE_TY, South_F, South_U), by = join_by(FEATURE_TY))
```

Add natural habitat as a polygon (based on AOI polygon)

```{r}
AOI_Feature <- AOI %>% 
  mutate(FEATURE_TY="NATURAL", SOURCE="AOI", YEAR=NA, HFI_ID=NA, YEAR_SOURCE=NA, 
         Shape_Length=NA, Shape_Area=NA, South_F=0, South_U=0) 
st_geometry(AOI_Feature) <- "geom"
Order <- Order <- match(names(HF2021_SE), names(AOI_Feature))
AOI_Feature <- AOI_Feature[,Order]
HF2021_SE <- rbind(HF2021_SE, AOI_Feature)
```

Create two new variables:

- H = mean(F, U):   combined deviation from naturalness
- Q = 1 - H:   degree of naturalness = habitat quality

```{r}
HF2021_SE <- HF2021_SE %>% mutate(H = (South_F+South_U)/2) %>% mutate(Q = 1 - H)
head(HF2021_SE)
```

## 3. Rasterize habitat quality Q and derive resistance values

```{r}
template <- terra::rast(terra::ext(vect(AOI)), res = 1000)
Q_stack <- terra::rast(fasterize(HF2021_SE, raster::raster(template), 
                                 field="Q", by="FEATURE_TY"))
```

We use the fuzzy algebraic sum to combine the Q values for mixed pixels (Marrec et al. 2020):

```{r}
Q <- prod(Q_stack, na.rm=TRUE)
#plot(Q)
```

Resistance values are now derived as a power function of H = 1 - Q (omitting the effects of water bodies and slope)

```{r}
R <- (2 - Q)^10
range(values(R), na.rm=TRUE)
```
conductance <- 1/R

The above function of Q can be supplied as the cost function in ConScape.

## 4. Aggregate rasters using mean

Here we use a factor of 4, to get 100 m resolution from rasters of 25 m

I swapped out "mean" for "max" to take a more conservative approach regarding resistance

```{r}
R.2000 <- terra::aggregate(R, fact=2, fun="max")
Q.2000 <- terra::aggregate(Q, fact=2, fun="mean")
```

```{r}
Rasters.1000 <- c(Q, R)
Rasters.2000 <- c(Q.2000, R.2000)
#Rasters.25 <- c(Q, R)
```

## 5. Export rasters

```{r}
writeRaster(Rasters.1000, here::here("data/Rasters.1000.tif"), overwrite=TRUE)
writeRaster(Rasters.2000, here::here("data/Rasters.2000.tif"), overwrite=TRUE)
```


